---
title: "HW 03 - Great Lakes Fish Datasets"
author: "Daytonna Anderson"
date: "October 8, 2025"
output: html_document
---

## Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse)
```

```{r load-data, message = FALSE}
catch_data <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-06-08/fishing.csv')

fish_stock <- read_csv("data/great-lakes-fish-stock.csv")
```

## Exercises

### Exercise 1

```{r catch-data-subset}
catch_data <- catch_data %>%
  select(year, lake, species, region, values)
```

```{r fish-stock-subset}
fish_stock <- fish_stock %>%
  select(YEAR, MONTH, LAKE, STATE_PROV, SPECIES, WEIGHT)
```

### Exercise 2

```{r new-glimpse}

glimpse(fish_stock)

glimpse(catch_data)
```

### Exercise 3

```{r simple-fish}
fish_stock <- fish_stock %>% 
  filter(STATE_PROV %in% c("MI", "WI", "IL", "IN"),
         SPECIES == "CHS",
         LAKE == "MI")
```

```{r simple-catch}
catch_data <- catch_data %>% 
  filter(region == "U.S. Total",
         species == "Chinook Salmon",
         lake == "Michigan")
```

### Exercise 4

```{r data-frame}
nrow(fish_stock)
nrow(catch_data)

min(fish_stock$YEAR)
max (fish_stock$YEAR)

min(catch_data$year)
max(catch_data$year)
```

There are less observations within the specific data frame that we set for one data set over the other.

### Exercise 5

```{r finalize-fish-stock, eval = TRUE}
fish_stock_yearly <- fish_stock_yearly %>% 
  rename(year = YEAR) %>% 
  mutate(stock_weight_yr_lbs = as.integer(STOCK_WEIGHT_YR * 0.454 / 1000)) %>% 
  select(year, stock_weight_yr_lbs)
```

```{r finalize-catch-data, eval = TRUE}
catch_data <- catch_data %>% 
  rename(catch_weight_yr_lbs = values)
```

### Exercise 6

Remove this text, and add your answer for Exercise 6 here.

### Exercise 7

Remove this text, and add your answer for Exercise 7 here.

### Exercise 8

Remove this text, and add your answer for Exercise 8 here.
